### 2016-05-16 text/Chapter04.tex
Edits in further work and the rest (Chapter 8)

In the next section, we move from abstract treatment of the flat coeffect calculus to a more
concrete discussion. We explain its category-theoretical motivation, we use it to define
translational semantics (akin to Haskell's do notation) and we prove a soundness result that
well-typed programs in flat coeffect calcului for implicit parameters and dataflow do not get
stuck in the translated version.

>>>>

In the next section, we move from abstract treatment of the flat coeffect calculus to a more
concrete discussion. We explain its category-theoretical motivation, we use it to define
translational semantics (akin to Haskell's ``do'' notation) and we prove a soundness result that
well-typed programs in flat coeffect calcului for implicit parameters and dataflow do not get
stuck in the translated version.

### 2016-05-16 text/Chapter05.tex
Edits in further work and the rest (Chapter 8)

The development in this chapter closely follows the example of effectful computations.
Effect systems provide a type system for tracking effects and monadic translation can be used as
a basis for implementing effectful domain-specific languages (e.g.~through the do-notation in
Haskell).

@@

\paragraph{Properties.}
The categorical semantics  in Section~\ref{sec:semantics-translation} defines a translation
that embeds context-dependent computations in a functional programming language, similarly to how
monads and the do notation provide a way of embedding effectful computations in Haskell.

@@

Although the notion of indexed comonads presented in the previous section is novel and interesting
in its own, the main reason for introducing it is that we can view it as a translation that provides
embedding of context-aware domain-specific languages in a simple target functional language.
In this section, we follow the example of effects and monads and we use the semantics to define a
translation akin to the do notation in Haskell.

@@

We then used the comonadic semantics to define a \emph{comonadically-inspired translation}
that turns programs written in a domain-specific coeffect language into a functional target
language. This is akin to the Haskell do-notation for monads. Finally, we extended the target
language with concrete implementations of comonadic operations for dataflow and implicit parameters
and we presented a syntactic safety proof. In summary, the proof states that well-typed
context-aware programs written in a coeffect language \emph{do not go wrong} (when translated
to a simple functional language and evaluated).

>>>>

The development in this chapter closely follows the example of effectful computations.
Effect systems provide a type system for tracking effects and monadic translation can be used as
a basis for implementing effectful domain-specific languages (e.g.~through the ``do'' notation in
Haskell).

@@

\paragraph{Properties.}
The categorical semantics  in Section~\ref{sec:semantics-translation} defines a translation
that embeds context-dependent computations in a functional programming language, similarly to how
monads and the ``do'' notation provide a way of embedding effectful computations in Haskell.

@@

Although the notion of indexed comonads presented in the previous section is novel and interesting
in its own, the main reason for introducing it is that we can view it as a translation that provides
embedding of context-aware domain-specific languages in a simple target functional language.
In this section, we follow the example of effects and monads and we use the semantics to define a
translation akin to the ``do'' notation in Haskell.

@@

We then used the comonadic semantics to define a \emph{comonadically-inspired translation}
that turns programs written in a domain-specific coeffect language into a functional target
language. This is akin to the Haskell ``do'' notation for monads. Finally, we extended the target
language with concrete implementations of comonadic operations for dataflow and implicit parameters
and we presented a syntactic safety proof. In summary, the proof states that well-typed
context-aware programs written in a coeffect language \emph{do not go wrong} (when translated
to a simple functional language and evaluated).

### 2016-05-16 text/Chapter07.tex
Edits in further work and the rest (Chapter 8)

\noindent
The list mirrors a list of steps that need to be done when supporting a new effectful computation
in a language that supports monadic do-notation. The step (3) corresponds to implementing a new
monad and (4) corresponds to adding monad-specific effectful operations. The step (2) applies when
using indexing to track effects more precisely \cite{effects-embedding}. The only step that does not
have a counterpart in effectful/monadic languages is (1).

@@

The last interactive element of the essay lets the reader explore the translation of source
context aware language to a target simple functional language (Section~\ref{sec:impl-theory-transl}).
Compared with the monadic do-notation \cite{other-haskell98}, the comonadic translation is more
complex for two reasons.

>>>>

\noindent
The list mirrors a list of steps that need to be done when supporting a new effectful computation
in a language that supports monadic ``do'' notation. The step (3) corresponds to implementing a new
monad and (4) corresponds to adding monad-specific effectful operations. The step (2) applies when
using indexing to track effects more precisely \cite{effects-embedding}. The only step that does not
have a counterpart in effectful/monadic languages is (1).

@@

The last interactive element of the essay lets the reader explore the translation of source
context aware language to a target simple functional language (Section~\ref{sec:impl-theory-transl}).
Compared with the monadic ``do'' notation \cite{other-haskell98}, the comonadic translation is more
complex for two reasons.
